\documentclass[10pt,aspectratio=169]{beamer}

% All the boilerplate is in ccaslides.sty
% Note that this also pulls in a custom vogtwidebar.sty
\usepackage{ccaslides}

\author{Ji\v{r}\'i Lebl}

\institute[OSU]{%
Departemento pri Matematiko de Oklahoma {\^S}tata Universitato}

\title{Cultivating Complex Analysis:\\%
Meromorphic functions (5.2.4)}

\date{}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\begin{definition}
A holomorphic function $f \colon U \setminus S \to \C$ with poles on a
discrete set $S \subset U$ is said to be \emph{meromorphic}.
\end{definition}

\pause

If $p \in S$ is a pole, set
$f(p) = \infty$ to get a function
\[
f \colon U \to \C_{\infty} .
\]

\pause
The extended $f$ is continuous.

\pause
\medskip

In fact, $\nicefrac{1}{f}$ is holomorphic at a pole $p$
\quad
(a reasonable definition of holomorphic).

\pause
\medskip

A \emph{meromorphic function} is
``a holomorphic function $f \colon U \to \C_{\infty}$.''

\pause
\medskip

\textbf{Technicality:} Should we consider the constant $\infty$
a meromorphic function?

\pause
In this course, we do not.

\pause
\medskip

To emphasize we will often say
``Let 
$f \colon U \to \C_\infty$ be meromorphic.''

\end{frame}

\begin{frame}

One could define functions on $U \subset \C_\infty$ (like we did with LFTs).

\pause
\medskip

If $\infty \in U$,
\quad
$f \colon U \to \C_\infty$ is holomorphic at $\infty$ if
$f(\nicefrac{1}{z})$ is holomorphic at $0$.

\pause
\medskip

So an LFT is a biholomorphic mapping $\C_\infty \to \C_\infty$.

\pause
\medskip

\textbf{Exercise:}
Show that a holomorphic $f \colon \C_\infty \to \C_\infty$ has
at most finitely many poles and finitely many zeros.

\pause
\medskip

\textbf{Exercise:}
Show that a holomorphic $f \colon \C_\infty \to \C_\infty$ is
either constant or onto.

\pause
\medskip

\textbf{Exercise:}
Show that a holomorphic $f \colon \C_\infty \to \C_\infty$ is a
rational function (a polynomial divided by a polynomial).

\pause
\medskip

\textbf{Exercise:}
Show that an injective holomorphic $f \colon \C_\infty \to \C_\infty$ is an
LFT\@.
\end{frame}

\end{document}
